---
title: "Problem Set 4"
author: "NAME HERE"
date: "9/17/2020"
output: pdf_document
---

Due date: Monday, September 28

For this problem set you will tidy up a dataset of 500 individuals. We also want to calculate each individual's BMI and appropriately categorize them.

Load your data:

```{r, include=FALSE}

```

### Question 1

Clean the column headers to have all lower case and no spaces and rename location_information to location.

```{r}

```

\newpage 

### Question 2

Create a new variable that calculates BMI for each individual.

You will need to navigate the different system of measurements (metric vs imperial). Only the United States is using imperial.

* BMI calculation and conversions:
    + metric:   $BMI = weight (kg) / [height (m)]^2$
    + imperial: $BMI = 703 * weight (lbs) / [height (in)]^2$
    + 1 foot = 12 inches
    + 1 cm = 0.01 meter

Although there's many ways you can accomplish this task, we want you to use an if_else(). 

```{r}

```

\newpage

### Question 3

Create a new variable that categorizes BMI with case_when():

* Underweight: Below 18.5
* Normal: 18.5-24.9
* Overweight: 25.0-29.9
* Obese: 30.0 and Above

```{r}

```

Could we have used if_else()?  


\newpage

### Question 4

Arrange your data by location and descending order of bmi.

```{r}

```

### Question 5

Filter your data to remove the height, weight, and bmi column.

```{r}
```

\newpage

### Challenge

Perform all the actions in this problem set with one dpylr call. 

```{r}

```
