---
title: "Problem Set 8"
author: "name?!"
date: "date?!"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

Due date: Wednesday, November 11th

## From Problem Set \#7, Part 2

If you previously completed this part please copy/paste your code below.

### Question 7 / Challenge

Use the readxl library and load two data sets from the
"two\_data\_sheets" file. There's a parameter that you can specify which
sheet to load with read\_excel(). In this case, we have data about rat
reaction time in sheet 1 and home visits in sheet 2.

```{r, echo=T, results="markup"}
library(tidyverse)
library(readxl)
df_rats <- read_excel("../data/two_data_sheets.xlsx", 1)
df_home <- read_excel("../data/two_data_sheets.xlsx", 2)
```

\newpage

### Question 8

For the rats data, pivot the data frame from wide to long format. We
want the 1, 2, 3 columns, which represent the amount of cheese placed in
a maze, to transform into a column called "cheese". The values in the
cheese column will be the time, which represents the amount of time the
rat took to complete the maze. Please use the head() function to print
your data frame.

```{r}
# convert from wide to long
df_rats$subject <- factor(df_rats$subject)
df_rats_long <- df_rats %>% 
  pivot_longer(c(`1`, `2`, `3`), names_to = "cheese", values_to = "time")

head(df_rats_long)
```

\newpage

### Question 9

Compute the mean and standard deviation of the maze time.

```{r}
df_rats_long %>%
  # organize by amount of cheese
  group_by(cheese) %>%
  # summarize
  summarize(mean = mean(time), # mean function
            sd = sd(time))     # standard deviation function
```

\newpage

### Question 10

With the home visits data, pivot the data frame
from long to wide. We want the names from the action column to become
unique columns and the values to represent the counts. Please use the
head() function to print your data frame.

```{r}
# pivot from long to wide
df_home_wide <- df_home %>%
  pivot_wider(names_from = action, values_from = count)

head(df_home_wide)
```

\newpage

## Problem Set 8

For this problem set we will work with four tables that are relational
to each other. The following keys link the tables to each other:

* patient_id: patients, schedule
* visit_id: schedule, visits
* doctor_id: visits, doctors

```{r, echo=F, include=F, results="markup"}
patients <- read_csv("../data/patients.csv")
visits <- read_csv("../data/visits.csv")
doctors <- read_csv("../data/doctors.csv")
schedule <- read_csv("../data/schedule.csv")
```

### Question 1

You've been tasked to collect information on patients within the patients
data set. To start this task, you need to join the patient data to the
schedule data. We only want to keep the observations that are both
present in the patient data AND the schedule data.

Which kind of join do you use?

**inner join**

How many observations do you see? Note: Some patients have multiple
visits.

**124 observations**

```{r}
# inner join by patient_id
inner.join.patient <- patients %>% 
  inner_join(schedule, by = "patient_id")

#str(inner.join.patient)
```

\newpage

### Question 2

In the visits data, we have a variable called "follow_up" where Y means
a follow-up is needed and N means a follow-up is not needed. How many of
these patients require a follow-up? You will want to first make a join
and then subset. Use the data frame created from the previous question.

Which join did you use?

**left join**

How many patients need a follow-up?

**27**

```{r}
left.follow.up <- inner.join.patient %>%
  left_join(visits, by = "visit_id")

# two ways we can filter:
follow.up <- left.follow.up %>% filter(follow_up == "Y")
follow.up <- left.follow.up[which(left.follow.up$follow_up == "Y"), ]
```

\newpage

### Question 3

Which doctors do these patients need follow-up with? Print out the
doctors. You can use unique(). Call this data frame **doctors.contact**.

Which join did you use?

**left join**

```{r}
doctors.contact <- follow.up %>%
  left_join(doctors, by = "doctor_id")

unique(doctors.contact$doctor)
```

\newpage

### Question 4 / Challenge

Find out which patients that are in schedule but not in doctors.contact.
Hint: anti_join()

Create a unique list of these patients with their demographic
information from the patients data.

```{r}
not.in.contact <- schedule %>%
  anti_join(doctors.contact, by = "patient_id") %>% # drop observations in schedule
  left_join(patients, by = "patient_id") %>% # join demographic information
  select(-c(visit_id, date)) %>%  # drop variables that prevent us from using distinct() properly
  distinct() # grab distinct rows

head(not.in.contact)
```
