---
title: "Problem Set 3"
author: "YOUR NAME HERE"
date: "DATE"
output: pdf_document
urlcolor: cyan
---

* Due date: Monday, September 21
* Submission process: Please submit your assignment directly to Gradescope. You can do this by knitting your file and downloading the PDF to your computer. Then navigate to Gradescope.com or via the link on BCourses to submit your assignment. 

Helpful hints:

- Knit your file early and often to minimize knitting errors! If you copy and paste code from the slides, you are bound to get an error that is hard to diagnose. Typing out the code is the way to smooth knitting. We recommend knitting your file each time after you write a few sentences/add a new code chunk, so you can detect the source of the knitting error more easily. This will save you and the teaching team time!

- Please make sure that your code does not run off the page of the knitted PDF. If it does, we can't see your work. To avoid this, have a look at your knitted PDF and ensure all the code fits in the file. When it doesn't, go back to your .Rmd file and add spaces (new lines) using the return or enter key so that the code runs onto the next line.

--------------------------------------------------------------------------------
```{r, include=FALSE}
# insert libraries that you will need
library(tidyverse)
library(lubridate)
```

\newpage 

Before you start: [Please make a conscious effort to write clear and readable code](https://jef.works/R-style-guide/) through this problem set (and your future R career). Adding comments for unclear code is a very good practice!

For this problem set we will use *fictional* data inspired by research on non-deceptive or open-label placebos. Non-deceptive placebos are placebos but without the deception. Some studies have found suggestions that, despite not being tricked, participants are reporting similar benefits to what they would have with placebos! You can read more here:

[NPR: Is A Placebo A Sham If You Know It's A Fake And It Still Works?](https://www.npr.org/sections/health-shots/2016/10/27/499475288/is-it-still-a-placebo-when-it-works-and-you-know-its-a-placebo)

[Nature Communications: Placebos without deception reduce self-report and neural measures of emotional distress](https://www.nature.com/articles/s41467-020-17654-y)

In this fictional data we conducted an experiment across two university sites to investigate whether non-deceptive placebos decreased self-report pain ratings. There were three groups: control, placebo, and non-deceptive placebo. Each participant completed a pre- and post- pain induction task and provided a pain rating. All participants completed the same procedures during the pre-test. Only during the post-test did participants in the intervention arms (placebo, non-deceptive) receive additional instructions prior to the pain induction task (i.e., placebo or non-deceptive placebo ratings).

Data coding:

- ID: Contains participant ID number, a letter to indicate group, and pre or post tags. 

  C = Control
  P = Placebo
  N = Non-deceptive
  
- LOCATION: Research Site

- PAIN RATING: Self report of pain based on a 0-10 scale

- DATE: Date of observation

### Question 1

Read in the data! To make it slightly more challenging we have changed the format from a .csv to .xlsx and "hidden" the data one level deeper in the /data folder. Take a look at the data to get oriented. Please use "df" as the name of your data frame for this problem set.

```{r}

```

\newpage

### Question 2

It's a bit difficult to tell what group each participant is in with their IDs combined with their grouping. Create a new column called "GROUP" based on the letter assignment for IDs. Makes sure to following naming convention above.

```{r}

```

\newpage 

### Question 3

We have a similar issue telling apart the pre- and post- observations. Create a new column called "TEST" that distinguishes whether the observation is a pre- or post-test.

Unfortunately, the two research sites were not consistent in their naming convention. You will need to consider the different cases. 

```{r}

```

\newpage 

### Question 4

Again, there were difference in the formatting for dates across the two research sites. Create a new column called "DATE_FIX" that grabs only the date. Make sure this new date column takes the following format: yyyy-mm-dd

Hint: Check out ?parse_date_time

```{r}

```

\newpage 

### Question 5

You realize there was a strange error in your excel file that pushed the years forward by 1 year and behind by 7 days. Rather than editing your excel sheet and potentially making an incorrect permanent change to your raw data you decide to fix the error in R. Create a new column called "DATE_FIX_2" that fixes the date.

```{r}

```

\newpage 

### Question 6

Let's clean up our data frame by removing DATE and DATE_FIX. Afterwards, rename DATE_FIX2 to DATE.

```{r}

```

\newpage 

### Question 7

We're interested in plotting our data to begin digging into the results. Below is dplyr and ggplot code that we will go over later in the course. For the time being run the following code. You may need to install ggthemes.

```{r}
# install.packages("ggthemes")
# library(ggthemes)

df_plot <- df %>% 
  group_by(GROUP, LOCATION) %>%
  summarize(MEAN_PAIN = mean(PAIN_RATE))

ggplot(df_plot, aes(x = LOCATION, y = MEAN_PAIN, fill = GROUP)) +
  geom_col(position = "dodge") +
  ylim(0, 10) +
  theme_few() +
  scale_fill_few("Medium") +
  theme(axis.title = element_blank(),
        axis.title.y = element_text()) +
  labs(fill = "Group",
       title = "Non-deceptive placebo study",
       y = "Pain rating")
```

For a quick first pass we think this visualization isn't so bad. However, logically, we think that the order of the groups should be: Control, Placebo, Non-deceptive. Make GROUP into a factor that reflects this order. If done correctly when you re-run the above chunk the plot should have the same order.

```{r}

```

\newpage 

### Question 8

We're interested in seeing whether pain ratings changed from pre to post. 

Create two list of the pain ratings. One list holds all of the pre-test while the other list holds the post-test.

Once you have these list create a third list called "diff" that calculates the difference between post and pre. 

```{r}


# uncomment after you create diff
# this code will create a table of counts
#table(diff)
```

\newpage

### Challenge

Take a look at this [rmarkdown reference guide](https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf).

On the first page are some very useful syntax to create *awesome* pdfs, htmls and, if we must, word documents. Practice a few of them below. The fun part (or frustrating part) is the repeated knitting to see if your intention worked out. 

Some suggestions: create bullet list, italics, bold, change the url colors, change font





You're done! Please knit to pdf and upload to gradescope. 