---
title: "Lab Session #1 9/22/20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)

```

## if...else statements

* create an if statement that assigns y a value of "greater than 10" when x>10
* create an if...else that does the above and assigns value of "less than or equal to 10" when x<=10
* create an if else that prints "West", "South", or "Everywhere else" based on state value
* try out the above with different values of x and state

```{r}

x <- 8


west_coast <- c("CA","AZ","WA")
south <- c("AZ","NM","TX")

state <- "CA"




```

## Importing a file

Import the following columns and ensure they are the corresponding data types:

* CDPHId - character
* BrandName - character
* PrimaryCategory - character 
* ChemicalName - character
* InitialDateReported - date
* ChemicalDateRemoved - character (will convert later)

```{r}

file_path <- "https://data.chhs.ca.gov/dataset/596b5eed-31de-4fd8-a645-249f3f9b19c4/resource/57da6c9a-41a7-44b0-ab8d-815ff2cd5913/download/cscpopendata.csv"

cos <- read_csv(
  file_path
)



```

## Structure and Milestone 1

* Structure of table
* Unique chemical names
* Frequency of product type
* Distribution of dates reported
```{r}



```

## Tidying the table

* get rid of duplicate lines
* keep only records that contain Titanium or Retinol in the name
* sort table to show by most recent date reported

```{r}


```

## Create new columns

* Convert ChemicalDateRemoved to date
  + Remove records with invalid ChemicalDateRemoved
* Has the chemical been removed
* Days to removal
* Days in product (if it has not been removed, used today as date)
* Create a removal category:
  + Removed before reported
  + Removed within 30 days of reporting
  + Removed within year of reporting
  + Not removed
    
```{r}



```

Return vector of Brands that were (a) reported in 2019-2020 and (b) removed Titanium dioxide within 30 days or initial report
```{r}

```




