---
title: "Midterm Solutions"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)

```

1. Which of the following will return a value of FALSE?
```{r}
x <- 5
y <- 25

x != y #TRUE
x*x == y #TRUE
x^2>y #FALSE
x<y #TRUE

```

2. For the following character string:

my_date <- “July 31, 2020”

Which code would yield a result five months prior to this date?

```{r}

my_date <- "July 31, 2020"

mdy(my_date) - months(5) #NA
mdy(my_date) %m-% months(5) #"2020-02-29"
as_date(my_date) %m-% months(5) #NA
as_date(my_date) - months(5) #NA

```

3. Which of these statements about vectors and lists is false?

b. A single vector can contain both numeric and character values

4. Which code will create a vector with the following contents?
2, 4, 6, 8, 10

```{r}
seq(2,10,by=2) #2  4  6  8 10 - CORRECT
2:10 #2  3  4  5  6  7  8  9 10
seq(2,10,length.out=4) #2.000000  4.666667  7.333333 10.000000
even(2,10) #Error

```

5. For the following vector:

v <- c(NA, 5, 10, NA, NaN)

What is the output of is.na(v)?
```{r}
v <- c(NA, 5, 10, NA, NaN)

is.na(v) #TRUE FALSE FALSE  TRUE  TRUE
```

6. For the list that is generated by this code:

multi_list<- list("Numbers" = seq(3,21,by=3), "Matrix" = matrix(c(-3,9,6,12,3,21), nrow = 2),"Words"=list("one","two","three"))

Which of the following will not return a single value of 3? 

```{r}
multi_list<- list("Numbers" = seq(3,21,by=3), "Matrix" = matrix(c(-3,9,6,12,3,21), nrow = 2),"Words"=list("one","two","three"))

 multi_list[["Numbers"]][1] #3
 multi_list[[2]][1,3] #3
 length(multi_list[["Words"]]) #3
 multi_list["Numbers"][1] #3  6  9 12 15 18 21 - CORRECT


```
7.  frames and tibbles are two options for storing tabular data in R. Which of the following is false? 

c. Tibbles cannot be indexed or subset in the same way as data frames (i.e. using [ ], [[]], or $)

8. Using the df below, which of the following will return a vector of the values of cases?
```{r}

df <- data.frame(
  state=c("CA","CA","AZ","AZ"),
  year=c(2019,2020,2019,2020),
  cases=c(34,23,89,27)
)

df$cases #vector
df[3] #df/tibble
df[[cases]] #error
df["cases"] #df/tibble

```

9. Using the df below, which of the following will not return this subset data frame (year 2020):
```{r}
 df[which(df$year==2020),]
 df[c(2,4)] #CORRECT (returns error)
 df[df$year==2020,]
 subset(df,year==2020)


```

10. There is a need for developing a function that calculates the volume of a rectangular shipping container  (length*width*height). If the volume is less than or equal to 1000 cubic feet, then the function should return “too small”, if it is greater than or equal to 2000 cubic feet it should return “too big”, and if it is between 1000-2000 cubic feet it should return “just right”.

Which function will return the correct value when 
Length (l)= 12
Width (w) = 5
Height (h)= 6

```{r}
check_volume_a <- function(l, w, h) {
  if(volume >= 2000) {
   return("too big")
  } else if (volume <= 1000){
    return("too small")
  } else {
    return("just right")
  }
}


check_volume_b <- function(l, w, h) {
  volume <- l*w*h
  
  if(volume >= 2000) {
   return("too big")
  } else if (volume < 2000){
    return("just right")
  } else {
    return("too small")
  }
}

check_volume_c <- function(l, w, h) {
  volume <- l*w*h
  
  if(volume >= 2000) {
   return("too big")
  } else if (volume <= 1000){
    return("too small")
  } else {
    return("just right")
  }
}

check_volume_d <- function(l, w, h) {
  volume <- l*w*h
  
  if(volume >= 2000) {
    output <- "too big"
  } else if (volume <= 1000){
    output <- "too small"
  } else {
    output <- "just right"
  }
  
  return(volume)
}


check_volume_a(12,5,6)
check_volume_b(12,5,6)
check_volume_c(12,5,6) #correct
check_volume_d(12,5,6)


```

11. Import the csv data file. (2 pts)
How many rows and columns are there?
Rows = 
Columns  = 
The “OSHPD Facility Number” column is reading in as character which is causing some ID’s to have leading zeros, creating some inconsistencies in the data. Force it to read in as numeric to drop the leading 0. Paste the argument used at import here:
Notice the column names are not reading in a very user-friendly way. Rename all columns to align with best practices for naming. Paste new column names here: 

```{r }
# payer <- read_csv(
#   "/home/rstudio/data/inpatient_payer_ca.csv",
#   col_types = "nncccnn"
# )

payer <- read_csv(
  "inpatient_payer_ca.csv",
  col_types = "nncccnn"
)

#Rows: 21899
#Cols: 7

payer <- payer %>%
 rename_with(~ tolower(gsub(" ","_",.x,fixed=TRUE)))

names(payer)

# few different ways so test out
```

12. The column that originally read in as “Expected Payer” contains categories that are inconsistently named, but obviously mean the same thing.  Clean up the values to include 9 categories. Paste the 9 unique categories here:

```{r}

unique(payer$expected_payer)

payer2 <- payer %>%
  mutate(expected_payer = gsub("_"," ", expected_payer))

#could also use a long case when, or other string functions

unique(payer2$expected_payer)
  
```
13. The values in the County column contain a mixture of upper case, lower case, and title case.  
Change the values to all be in a consistent case. 
Enter the number of unique county values before changes here:
Enter the number of unique county values after changes here:

```{r}
length(unique(payer2$county_name))
#182

payer3 <- payer2 %>%
  mutate(county_name = str_to_upper(county_name))

length(unique(payer3$county_name))
#56

```

14. [1 pt] There are some facilities with duplicate records for a year and payer type; in some cases the counts on these records differ. Retain only one record per year, facility id, and payer type; for instances where there are records with different counts, retain the record with the highest number.
How many rows remain?
What is the count of Private Coverage visits at Kern Medical Center for 2013?

```{r}

payer4 <- payer3 %>%
  arrange(year,oshpd_facility_number, expected_payer, desc(count)) %>%
  distinct(year,oshpd_facility_number, expected_payer, .keep_all=TRUE)

#21632
#733

```

15. [1 pt.] There are some facilities that have used negative numeric values to indicate unknown for the “Counts” column; these values are invalid. Replace any invalid values with missing. 
After replacing these values, how many rows are missing a value for “Counts”?
```{r}

payer5 <- payer4 %>%
  mutate(count = if_else(count<0, NA_real_, count))

sum(is.na(payer5$count))

#1200

```

16. Limit dataset to only contain rows for Medi-Cal coverage and years 2010-2015. Create a new column called pct_medi_cal that calculates the percent of inpatient stays that were covered by Medi-Cal. Display the percentage as multiplied by 100 and rounded to 1 decimal (for example, 35.1% as 35.1); replace any missing or invalid values for count or pct_medical with 0.
How many records are in the subsetted dataset?
What is the value of pct_medi_cal for Eden Medical Center in 2015?
```{r}
payer6 <- payer5 %>%
  filter(expected_payer=="Medi-Cal") %>%
  mutate(pct_medi_cal = round(100*count/total_inpatient_stays,1)) %>%
  replace_na(list(count=0,pct_medi_cal=0))

#2557
#28.4%

```

17. [1 pt] Create a new column called quartiles that categorizes percent based on quartile of percent paid by Medi-Cal. 
What is the lower bound for the top quartile?
What quartile is Woodland Memorial Hospital in for 2010?
```{r}

summary(payer6$pct_medi_cal)

payer7 <- payer6 %>%
  mutate(quartile = case_when(
    pct_medi_cal>38.80 ~ "q4",
    pct_medi_cal>23.75 ~ "q3",
    pct_medi_cal>10.20 ~ "q2",
    TRUE ~ "q1"
  ))

payer7 %>% filter(year==2010 & facility_name=="WOODLAND MEMORIAL HOSPITAL") %>% select(quartile)

#39.10
#Q3

```

18. [1 pt] Order the table to display rows by most recent year and percent of inpatient visits that were covered by Medi-Cal.
Enter the Facility Name with the highest proportion of inpatient visits covered by Medi-Cal in 2015. 
```{r}

payer8 <- payer7 %>%
  arrange(desc(year),desc(pct_medi_cal))

#	STAR VIEW ADOLESCENT - P H F
# CRESTWOOD PSYCHIATRIC HEALTH FACILITY-CARMICHAEL	
# CRESTWOOD PSYCHIATRIC HEALTH FACILITY-SACRAMENTO


```

19. [1 pt] There is interest in knowing specifically about payers for Kaiser facilities. Create a new variable to flag observations for Kaiser facilities. Restrict the dataset to only include records for Kaiser, Medi-Cal, Counts >0, and year is 2015. 
How many records remain?
What Kaiser facility has the highest percent of Medi-Cal covered inpatient stays?
```{r}
payer9 <- payer8 %>%
  mutate(kaiser_flag = if_else(str_detect(facility_name,"KAISER")==TRUE,"kaiser","not kaiser")) %>%
  filter(kaiser_flag=="kaiser" & count>0 & year==2015)

#32
#	KAISER FND HOSP - SOUTH SACRAMENTO

```


[Extra credit 2 pts] Using the dataset created in question #18, use dplyr functions to answer the following questions:
What county has the highest total number of Medi-Cal covered inpatient stays in 2015?
In 2014, which county had the highest mean facility-level percent of inpatient stays covered by Medi-Cal?


```{r}

payer_extra1 <- payer8 %>%
  filter(year==2015) %>%
  group_by(year, county_name, expected_payer) %>%
  summarise(county_total = sum(count)) %>%
  arrange(desc(county_total))

payer_extra2<- payer8 %>%
  filter(year==2014) %>%
  group_by(year, county_name, expected_payer) %>%
  summarise(mean_pct = mean(pct_medi_cal)) %>%
  arrange(desc(mean_pct))

#LOS ANGELES
#MADERA
  

```
