---
title: "Problem Set 7"
author: "name?!"
date: "date?!"
output: pdf_document
---

Due date: Monday, October 26th

```{r, include=F}
# load libraries and suppress output w/ include=F
library(tidyverse)
library(lubridate)
library(readxl)
```

## Part 1

For part 1 of this problem set we will work with motor vehicle crash data from New York City. You can read more about this [publicly available data set on their website](https://data.cityofnewyork.us/Public-Safety/Motor-Vehicle-Collisions-Crashes/h9gi-nx95).

The data is called "Motor_Vehicle_Collisions_Crashes". We want you to perform the following:

1. Rename the column names to lower-case and replace spaces with an underscore.
2. Select only:
   + crash_date
   + number_of_persons_injured
   + contributing_factor_vehicle_1
   + vehicle_type_code_1
3. Drop all rows with an NA value
4. Lower case the vehicle_type_code_1 variable and replace spaces with a dash.
5. Filter the data for vehicles that have a count/appear in the data set 500 times or more
   + Hints: group_by(), mutate(), n(), filter()
6. Calculate the percentage by vehicle
7. Which vehicle group accounted for 1.5% (0.0155) of the accidents?

We have grouped the questions below to push you to perform commands with less code. As you're building your code we recommend going line by line to test, then combining.

\newpage

### Questions 1-3

```{r, include=F}
# load your data
df_motor <- read_csv("../data/Motor_Vehicle_Collisions_Crashes.csv")
```


```{r}
df_motor <- df_motor %>%
  # lower case and remove spaces
  rename_with(~ tolower(gsub(" ","_", .x, fixed=TRUE))) %>%
  # select certain columns
  select(crash_date, crash_time, 
         number_of_persons_injured, 
         contributing_factor_vehicle_1, 
         vehicle_type_code_1) %>% 
  # drop NA rows
  drop_na()

dim(df_motor)
head(df_motor)
```

\newpage

### Questions 4-5

```{r}
# lower case vehicles and add dash between spaces
df_motor <- df_motor %>%
  mutate(vehicle_type_code_1 = 
           gsub(" ", "-", ignore.case=T, tolower(vehicle_type_code_1))) %>%
  # organize by vehicles
  group_by(vehicle_type_code_1) %>%
  # create a variable for counts
  mutate(count = n()) %>%
  # filter counts > 500
  filter(count > 500)
```

\newpage

### Question 6

```{r}
# calculate percentage by vehicle
df_motor %>%
  group_by(vehicle_type_code_1) %>%
  summarize(count = n(),
            perc = count/nrow(df_motor)) %>%
  arrange(perc)
```

Buses account for 1.5% of the accidents.

\newpage

### Question 7

Buses:

* count: 2862
* perc: 0.0155
